---
title: "Unsupervised learning -R"
author: "Mokaya Lydia"
output: html_document 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Defining The Question
The brand’s Sales and Marketing team of Kira plastinina would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.Perform clustering stating insights drawn from your analysis and visualizations. Provide comparisons between the approaches learned this week i.e. K-Means clustering vs Hierarchical clustering highlighting the strengths and limitations of each approach in the context of your analysis. 

# Understanding The context
Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia.

# Metric for Success
Our metric for success will be fairly accurate unsupervised model fitted to the data so as to cluster it into categories.
# Experimental Design 
1. Problem Definition
2. Data Sourcing
3. Check the Data
4. Perform Data Cleaning
5. Perform Exploratory Data Analysis  (Univariate, Bivariate & Multivariate)
6. Implement the Solution
7. Challenge the Solution
8. Follow up Questions

# Data Relevance

The dataset consists of 10 numerical and 8 categorical attributes. The 'Revenue' attribute can be used as the class label.
"Administrative", "Administrative Duration", "Informational", "Informational Duration", "Product Related" and "Product Related Duration" represents the number of different types of pages visited by the visitor in that session and total time spent in each of these page categories. The values of these features are derived from the URL information of the pages visited by the user and updated in real-time when a user takes an action, e.g. moving from one page to another. 
The "Bounce Rate", "Exit Rate" and "Page Value" features represent the metrics measured by "Google Analytics" for each page in the e-commerce site. 
The value of the "Bounce Rate" feature for a web page refers to the percentage of visitors who enter the site from that page and then leave ("bounce") without triggering any other requests to the analytics server during that session. 
The value of the "Exit Rate" feature for a specific web page is calculated as for all pageviews to the page, the percentage that was the last in the session.
The "Page Value" feature represents the average value for a web page that a user visited before completing an e-commerce transaction. 
The "Special Day" feature indicates the closeness of the site visiting time to a specific special day (e.g. Mother’s Day, Valentine's Day) in which the sessions are more likely to be finalized with the transaction. The value of this attribute is determined by considering the dynamics of e-commerce such as the duration between the order date and delivery date. For example, for Valentina’s day, this value takes a nonzero value between February 2 and February 12, zero before and after this date unless it is close to another special day, and its maximum value of 1 on February 8. 
The dataset also includes the operating system, browser, region, traffic type, visitor type as returning or new visitor, a Boolean value indicating whether the date of the visit is weekend, and month of the year.
```{r}
#installing the necessary libraries
install.packages("magrittr") # package installations are only needed the first time you use it
install.packages("dplyr")    # alternative installation of the %>%
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr) 
```
# Data Sourcing
```{r}
shoppers <- read.csv("online_shoppers_intention.csv")
head(shoppers)
```
# Check the Data
```{r}
#summary of the dataset
summary(shoppers)

```
```{r}
# information about the dataset
str(shoppers)
```

#Data Cleaning
### Completeness
```{r}
#Checking for Missing Values
colSums(is.na(shoppers))
```
```{R}

# Filling the null columns with their respective mean

shoppers  = shoppers %>%

    mutate(Administrative =replace(Administrative,is.na(Administrative),mean(Administrative,na.rm=TRUE)))%>%
    mutate(Administrative_Duration =replace(Administrative_Duration,is.na(Administrative_Duration),mean(Administrative_Duration,na.rm=TRUE)))%>%
    mutate(Informational = replace(Informational, is.na(Informational), mean(Informational, na.rm = TRUE)))%>%
    mutate(Informational_Duration =replace(Informational_Duration,is.na(Informational_Duration),mean(Informational_Duration,na.rm=TRUE)))%>%
    mutate(ProductRelated =replace(ProductRelated,is.na(ProductRelated),mean(ProductRelated,na.rm=TRUE)))%>%
    mutate(ProductRelated_Duration = replace(ProductRelated_Duration, is.na(ProductRelated_Duration), mean(ProductRelated_Duration, na.rm = TRUE)))%>%
    mutate(BounceRates =replace(BounceRates, is.na(BounceRates),mean(BounceRates,na.rm=TRUE)))%>%
    mutate(ExitRates = replace(ExitRates, is.na(ExitRates), mean(ExitRates, na.rm = TRUE)))

```

```{r}
# Confirming that we have no null values
sum(colSums(is.na(shoppers)))
```
## Consitency

```{r}
# Checking for Duplicates
sum(duplicated(shoppers))
```
```{r}
# Removing duplicates
# pick the non-duplicated rows
data = shoppers[!duplicated(shoppers), ]
dim(data)
```
```{r}
# Confirming if Duplicates are still there
sum(duplicated(data))
```
```{r}
# plotting a boxplot for the product related duration outliers

boxplot(data$Administrative,
main = "Administrative Boxplot",
xlab = "Administrative",
col = "orange",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for the Administrative_Duration
boxplot(data$Administrative_Duration,
main = "Administrative_Duration Boxplot",
xlab = "Administrative_Duration",
col = "blue",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for Informational

boxplot(data$Informational,
main = "Informational Boxplot",
xlab = "Informational",
col = "green",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for the Informational_Duration
boxplot(data$Informational_Duration,
main = "Informational_Duration boxplot",
xlab = "Informational_Duration",
col = "gold",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for the product related

boxplot(data$ProductRelated,
main = "Product Related Boxplot",
xlab = "Product Related",
col = "green",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```



```{r}
# plotting a boxplot for the product related duration 

boxplot(data$ProductRelated_Duration,
main = "Product Related Duration Boxplot",
xlab = "Product Related Duration",
col = "gold",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for the BounceRates

boxplot(data$BounceRates,
main = "BounceRates Boxplot",
xlab = "BounceRates",
col = "cyan",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for theExitRates

boxplot(data$ExitRates,
main = "ExitRates Boxplot",
xlab = "ExitRates",
col = "pink",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
```{r}
# plotting a boxplot for PageValues

boxplot(data$PageValues,
main = "PageValues Boxplot",
xlab = "PageValues",
col = "maroon",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
# Univariate Analysis

```{r}
# statistical description of the dataset
summary(data)
```
```{r}
# Histogram 
hist(data$Administrative)
hist(data$Informational)
hist(data$ProductRelated)
hist(data$BounceRates)
hist(data$ExitRates)
hist(data$TrafficType)
```
Bar plots
```{r}
# Bar plots of the categorical/factor modes variables
par(mfrow=c(4,1))
for(i in 11:16) {
	counts <- table(data[,i])
	name <- names(data)[i]
	barplot(counts, main=name, col = heat.colors(20))}
```
# Bivariate analysis

```{r}
# Plotting a scatter plot using the plot() method

plot(ExitRates ~ BounceRates, data = data, 
      col = "black",
      main = "Bounce vs Exit Rates Scatter Plot")

```
From the scatter plot there is a strong positive correlation between Exit rates and Bounce rates.
```{r}
# Scatter plot using ggplots and fitting a line of best fit

ggplot(data, aes(x = Administrative_Duration, y = Informational_Duration)) + 
        geom_point(size = 2, color= "green", shape = 23)+ 
        geom_smooth(method=lm,  linetype="dashed",color="darkred", fill="blue")+
        labs(title = "Info Duration vs Adm Duration Scatter Plot")
```
The is a positive correlation between the two variables but not very strong.

```{r}
# Stacked bar chart: Visitor Type vs Month
data %>%
    ggplot(aes(Month)) +
    geom_bar(aes(fill = VisitorType))+
    labs(title = "Stacked Chart: Visitor Type by Month")
```
observations

May, Nov, March, and December in that order are the busy months.
During these months there is a higher number of new visitors which the company can attract using offers tailored for them to retain them.
Feb and June are the least busy months.
```{r}
# Stacked bar chart: Revenue vs Month
data %>%
    ggplot(aes(Revenue)) +
    geom_bar(aes(fill = Month))+
    labs(title = "Stacked Chart: Revenue by Month")

```
Nov, May and March are best months when the company makes most revenue.




```{r}
# Stacked bar chart: Revenue vs Day Type
data %>%
    ggplot(aes(Revenue)) +
    geom_bar(aes(fill = Weekend))+
    labs(title = "Stacked Chart: Revenue by Day Type")

```
About three quarters of the data shows that a visit to the page did not result to the company making revenue i.e. the customer did not make a purchase.
Of the quarter remaining, the company made revenue mostly during the weekdays.

##correlation plot
###install packages
```{r, echo=TRUE, eval=TRUE}
install.packages("Hmisc")

install.packages("corrplot")

library("corrplot")
#plot the correlation table
cor <- cor(data[,1:10])
corrplot(cor, method="color")
```
administrative,administrative ,informational, informational duration, product related, product duration are all strongly positively correlated.
bounce rates and exit rates  also have a strong positive correlation.
```{r}
# pairplots
pairs(data[,c(1,2,3,4,6,7,8,9,10)])
```
# Multivarite analysis
```{r}
#PCA
data.pca <- prcomp(data[,c(1,2,3,4,5,6,7,8,9,10)], center = TRUE,scale. = TRUE)

summary(data.pca)
```
```{R}
# Visualize PCA
install.packages("DataExplorer")
library(DataExplorer)
# Subset your data to numerical columns only
num <- data[, c(1,2,3,4,5,6,7,8,9,10)]
plot_prcomp(num, variance_cap = 0.9, nrow = 2L, ncol = 2L)
```


# IMPLEMENTING THE SOLUTION
K-MEANS CLUSTERING
###changing our factor variables to numeric
```{r}
b = c("OperatingSystems", "Browser", "Region", "TrafficType")
for (i in b) {
    data[, i] = as.factor(data[, i])
}
```

```{r}
g = c('Month','VisitorType', 'Weekend')
for (i in g){
    data[,i] = as.numeric(data[,i])
}

```

### data preprocessing

```{R}

kdata<- data[, c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)]
data.class<- data[, "Revenue"]
head(kdata)
```
```{R}
# Normalize the data
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}

n = c(1,2,3,4,5,6,7,8,9,10)
for (i in n){
    kdata[,i] = normalize(kdata[,i])
}

head(kdata)

```
```{r}
# Applying the K-means clustering algorithm with no. of centroids(k)=3
# ---
# 
result <- kmeans(kdata,2)

```

```{r}
# Previewing the no. of records in each cluster
# 
result$size
```
```{r}
result$centers
```
```{r}
result$cluster
```
```{r}
k_data <- as.numeric(data.class)
table(result$cluster, k_data)
```
```{r}
mean(k_data == result$cluster)
```
HIERACHICAL CLUSTERING

```{r}
# Calculating euclidean distances of the independent variables.

d <- dist(kdata, method = "euclidean")

```

```{r}
# using the hclust clustering method.
h_model <- hclust(d, method = "ward.D2" )
```

```{r}
# plotting the dendogram
plot(h_model, cex = 0.6, hang = -1)

```
```{r}
# Cut tree into 2 groups.
sub_group = cutree (h_model, k = 2)
table (sub_group)
```
```{r}
table (sub_group, data.class)
```
```{r}
#checking accuracy
mean(sub_group == data.class)
```
# Challenging the solution
DBSCAN CLUSTERING
```{R}
install.packages('dbscan')
```
```{r}
# Loading the required library
library("dbscan")

# Applying our DBSCAN algorithm
# using a minimum of 4 points with in a distance of eps(0.4)
# 
l = c('OperatingSystems','Browser','Region','TrafficType')
for (i in l){
    kdata[,i] = as.numeric(kdata[,i])
}

db_model <- dbscan(kdata,eps=0.4,MinPts = 4)

```
```{r}
print(db_model)
```
```{r}
hullplot(kdata,db_model$cluster)
```







